---
title: "Introduction to statistics"
subtitle: "with applications in R"
author: "Dr. Jigme Tenzin"
institute: "College of Natural Resources"
date: "13 January 2021"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    css:
      - default
      - css/cnr.css
      - css/cnr-fonts.css
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
    includes:
      after_body: [css/insert-logo.html]
---


```{r setup, include = FALSE}
library(knitr)
library(ggplot2)
library(tidyverse)
library(nhsrtheme)
library(xaringanExtra)
library(icon)
xaringanExtra::use_tile_view()
xaringanExtra::use_share_again()
xaringanExtra::use_panelset()
xaringanExtra::use_clipboard()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,
  mute_unhighlighted_code = TRUE
)
style_panelset_tabs(foreground = "honeydew", background = "seagreen")
xaringanExtra::use_editable(expires = 1)
# set default options
opts_chunk$set(echo = TRUE,
               fig.width = 7.252,
               fig.height = 4,
               dpi = 300
               )

# uncomment the following lines if you want to use the NHS-R theme colours by default
# scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
# scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
# scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
# scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)
```

class: title-slide, left, bottom

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$date`


---
class: center, top
# About me

<img src ="img/jigme.jpg" width= "200", border-radius=50%>

### Dr. Jigme Tenzin

Department of Animal Science <br>
College of Natural Resources <br>
Royal University of Bhutan <br>
```{r echo=FALSE, fig.show='asis'}
icon::fontawesome('twitter')
```
[tenz_jigme](http://twitter.com/tenz_jigme)
```{r echo=FALSE, fig.show='asis'}
icon::fontawesome('envelope')
```
[jtenzin.cnr@rub.edu.bt](mailto:jtenzin.cnr@rub.edu.bt)
```{r echo=FALSE, fig.show='asis'}
icon::fontawesome('telegram')
```
[jigme77](https://t.me/Jigme77)

---
class: left, top

# Definition of terms 
----

--
### Statistics

Collect, analyze, summarize, interpret, and draw conclusions from the data
or actual numeric descriptions of sample data 

--
### Data
Any observations that you have collected or informatin gathered

--
### Population
The complete set of elements that are being studied or a particular group of interest

???
Population example: 
All males in the world.
All females in the world.
All children between 6 - 9 years of age.

--
### Samples
Some subset of a population or subset of population from which data is collected

--
### Census
Collecting from every member of a population

---
class: left, top

# Definiton of terms
----
.panelset[
.panel[.panel-name[Parameter]
### Parameter
A characteristics of a population (*mnemonic*: P-P) or numerical description of a poulation characteristics.
#### Examples
 - Mean height of all males in the world
 - Mean IQ of all females in Bhutan
 - 75% of all kids ages between 6 - 9 play video games
]

.panel[.panel-name[Statistics]
### Statistics
A characteristics of a sample (*mnemonic*: S-S) or numerical description of sample characteristics

#### Examples
- of 100 female asked, 47% dislike chocolate.
]
]

---
class: left, top

## Branches of statistics

.panelset[
.panel[.panel-name[Descriptive statistics]
Gather, sort, and summarize data from samples.

#### Examples
- 65% of seniors in local high school applying to college plan to major in business
]

.panel[.panel-name[Inferential statistics]
It uses descriptive statistics (data) to estimate population parameters.

#### Examples
- Based on phone survey, 22% of all men don't like football
]

]

---
class: left

# Applying definitions in statistics
 
--
#### The height of every fourth bottle in assembly line. Sample or population?

--
.can-edit[
 - Type the answer here]

--
#### Ages of all USA presidents. Population or sample?

--
.can-edit[
- Type the answer here]

--
#### A researcher stops 100 people in a store to survey household income. Identify population and the sample.

--
.can-edit[
* Population: 
* Sample:
]

--
#### Average number of hours per week a sample of 10 year olds spends watching is 20 hours. What is it represent: parameter or statistics? 

--
.can-edit[
* Type your answer here
]

--
#### 87% of all patients in a hospital report having an alcohol problem. Is it a parameter or statistic?

--
.can-edit[
* Type your answer here
]


---
class: left, top

# Types of data
----
.panelset[
.panel[.panel-name[Categorical variable]
Also known as class variable, or qualitative variable or nominal variables

#### Examples
- Sex
- Genotype
- Eye color (blue vs brown vs other)
- Drug treatment ( aspirin vs ibuprofen)
- Province
- Survival (live or die)

]

.panel[.panel-name[Numerical Variable]
Also known as quantitative variable. It can be either discrete or continuous

#### Examples
- Height 
- Weight 
- Tail length
- Dose (microgram per ml)
- Longevity (number of years)

]

.panel[.panel-name[Discrete]
Can be counted (Indivisible units)

#### Examples
- Number of limbs
- Litter size
- Number of offspring
- Number of petals
]

.panel[.panel-name[Continuous]
Can be measured (Infinitely divisible units)

#### Examples
- Arm length
- Height
- Weight
- Age
- Temperature 
- Duration of time
]

.panel[.panel-name[Measurement Scale]
.pull-left[## Categorical variable

### Nominal scale
eg. species, sex, diet
* No ordering 

### Oridinal scale
eg. small/medium/large
* Rank order 
* Differences and ratio undefined or only approximate
]

.pull-right[
## Numerical variable

### Interval scale 
eg. Celsius temperature, Years BC
* Arbitrary zero point
* Differences defined, ratios undefined

### Ratio scale
eg. mass, length, abundance, duration

* Physically meaningful zero point
* Differences and ratios defined

]

]
]

---
class: top, left
----

# Experiment and observational studies

.left-column[
## Experiment
- Measures specific traits after applying a treatment

## Observational studies
- Measures specific traits but doesn't modify subjects
]

.right-column[
## Sampling
Random - each member in a population has equal chance of being in that sample.
1. Simple random sample: each group of size 'n' has equal chance of being in that sample.

2. Convenience sampling: use the results you easily. Not random, not used.

3. Stratified random sampling: break population into sub-groups based on some characteristics, then take simple random sample of each subgroup.

4. Cluster sample: break population into 'cluster,' regardless of characteristics, randomly select certain number of clusters and then collect data from cluster

Strata: Homogeneous group

Cluster: Heterogenous group

]


---
class: left, top
# Introduction to R
----


.pull-left[
[R installation can be done from here](https://cran.r-project.org/)

### It is like a giant calculator
```{r basic-calc, echo=TRUE}
2 + 2 # Addition
2 - 4 # Subtraction
3 * 2 # Multiplication
```

]

.pull-right[
## Why use R?
### Advantages of R
- Free and open source
- High quality plots
- Reproducible reports 
- Continuously growing
- Many packages

### Disadvantages 
- Complicated language
- Spread across various packages
- Inefficient memory allocation
]

---
class: center, left
# Histogram 
----
.pull-left[
```{r first-plot, fig.height=5, echo=FALSE}
hist(mpg$hwy, 
     xlab = "Highway mileage", 
     ylab = "Number of cars", 
     main = "Histogram of mileage on Highway for different models of car")
```
]

.pull-right[
```{r, ref.label = 'first-plot', echo=TRUE, fig.show ='hide'}
```
]

## "Can use minimal codes to produce publication quality charts."





